\chapter*{Acknowledgements}

This dissertation would not have been possible without the support of the following people.

To my supervisors Alan and John for their patience, guidance and sharing of knowledge during this process. Thank you, it has been a real pleasure working together.

To my lecturers at the Waterford Institute of Technology, you have all been so open and free with your sharing of knowledge, it has been an absolute pleasure learning from you.

To my partner Michael, I'm keenly aware standing in the freezing cold at 02:00 helping me put up radio antennas might not have been the most exciting thing you've partaken in all year, but I couldn't have done it without your help. Thanks for your constant encouragement and for putting up with me during this. 

To my sister Jean, you inspired me to work as hard as I have these last 6 years. I wanted to make you as proud of me as I am of you. If you beat me to Ph.D, I'll buy you a bag of chips in Roccos.

To my father, I can only imagine the conversations we might have had, I suspect you would have loved playing with this equipment as much as I have.

To my mother, thank you for your friendship. You have always been a positive influence in my life. Some day you must sit down and explain how you can walk past, and in the time it takes to cross a room and without any prior involvement, solve a problem I have been struggling with for days.

To my four dogs Noah, Oscar, Rosie and Dora, you are the best mutts a man could ask for.